


# Catch up Clustering Day

## Learning Goals {-}


- Implement k-means and hierarchical clustering and interpret in their outputs and algorithms
- Synthesize and apply concepts covered so far on real data


<br><br><br>

## Real Data Clustering {-}

As a group, choose one of the following three datasets to work with:

1. Wine Attributes (download [here](template_rmds/wine.csv))
- 178 Italian wines were analyzed
- Variables (from Chemical Analysis)
  - Alcohol
  - Malic acid
  - Ash
  - Alcalinity of ash
  - Magnesium
  - Total phenols
  - Flavanoids
  - Nonflavanoid phenols
  - Proanthocyanins
  - Color intensity
  - Hue
  - OD280/OD315 of diluted wines
  - Proline

```{r}
library(readr)
wine <- read_csv('wine.csv')
```

2. Mall Customers (download [here](template_rmds/mall_customers.csv))
- 200 individuals
- Variables
  - Binary Gender
  - Age
  - Annual Income (in $1000's)
  - Spending Score (summary of buying behavior)
  
```{r}
library(readr)
customers <- read_csv('mall_customers.csv')
```  
  
3. Credit Card Clients (download [here](template_rmds/creditcard.csv))
- Almost 9000 credit card holders
- Variables based on 6 months of time
  - CUSTID: Identification of Credit Card holder
  - BALANCE : Balance amount left in their account to make purchases
  - BALANCEFREQUENCY : How frequently the Balance is updated, score between 0 and 1 (1 = frequently updated, 0 = not frequently updated)
  - PURCHASES : Number of purchases made from account
  - ONEOFFPURCHASES : Maximum purchase amount done in one-go
  - INSTALLMENTSPURCHASES : Amount of purchase done in installment
  - CASHADVANCE : Cash in advance given by the user
  - PURCHASESFREQUENCY : How frequently the Purchases are being made, score between 0 and 1 (1 = frequently purchased, 0 = not frequently purchased)
  - ONEOFFPURCHASESFREQUENCY : How frequently Purchases are happening in one-go (1 = frequently purchased, 0 = not frequently purchased)
  - PURCHASESINSTALLMENTSFREQUENCY : How frequently purchases in installments are being done (1 = frequently done, 0 = not frequently done)
  - CASHADVANCEFREQUENCY : How frequently the cash in advance being paid
  - CASHADVANCETRX : Number of Transactions made with "Cash in Advanced"
  - PURCHASESTRX : Number of purchase transactions made
  - CREDITLIMIT : Limit of Credit Card for user
  - PAYMENTS : Amount of Payment done by user
  - MINIMUM_PAYMENTS : Minimum amount of payments made by user
  - PRCFULLPAYMENT : Percent of full payment paid by user
  - TENURE : Tenure of credit card service for user

```{r}
library(readr)
credit <- read_csv('creditcard.csv')
```

### Your Goal {-}

- Goal: Cluster the data to discover insight and patterns in the data

- Available Methods
  - K-means with all quantitative variables
  - Partitioning around Medoids (pam) as a robust version of K-means
    - If you have at least one categorical variable, `daisy()` will calculate [Gower's distance](https://stat.ethz.ch/R-manual/R-devel/library/cluster/html/daisy.html)
  - Hierarchical clustering
    - If you have at least one categorical variable, `daisy()` will calculate [Gower's distance](https://stat.ethz.ch/R-manual/R-devel/library/cluster/html/daisy.html)

```{r}
kmeans(data, centers = k)

library(cluster)
pam(daisy(data), k = k)

hclust(daisy(data))
```

- Insights
  - Based on clustering, you'll want to interpret/visualize the resulting clusters to gain insight

- Deliverable
  - Create 1 graphic that demonstrates your insight
  - Add it to this [Shared Folder](https://drive.google.com/drive/folders/15O-WcCUBfvDYY-C5tb4ppFzQRhMfAQLc?usp=sharing) with your names in the filename


## Clustering in the Wild {-}

To give you a taste of how these methods get use in "the wild" world of science, here are a few papers (quality varies):

- Image Segmentation (https://link.springer.com/article/10.1007/s11042-021-10594-9)
- Bacteria Clustering (https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0002843)
- Document Clustering (https://www.ncbi.nlm.nih.gov/pmc/articles/PMC7790388/)
- Clustering ICD10 Diagnosis Codes (https://arxiv.org/abs/1909.00306)
- Clustering Activity Sequences (https://www.sciencedirect.com/science/article/abs/pii/S0968090X21000395?via%3Dihub)
